Fish do not need visas: the transboundary nature of the world’s exploited marine fish species (79/96 characters)

### 2020-02-04 Manuscript_Draft.Rmd
First corrections on comments from co authors

Forty years after the formal adoption of UNCLOS, accurate estimates of the number of exploited marine species that are shared by neighboring nations (hereafter called transboundary species [@Munro:2002uf]) is still unclear due to data limitations [@Caddy:1997ue], and methodological approaches [@Teh:2015gd]. In here, we aimed to estimate the number of exploited marine fish species shared by neighbouring countries and determine their contribution to global catch and fisheries revenue.

We overlay the known distribution of **968** commercially valuable marine species and the EEZ of 153 coastal countries and developed a four-criteria methodology to determine if a species was transboundary. The first criteria consisted in only considering shared species between neighboring EEZs, rather than the specie's extended distribution. For the second criteria we considered a species to be present in a grid cell if all data sources predicted positive occurrence. The third criteria determined whether the species were exploited at the grid cell using spatially explicit catch data. A species was kept in a grid cell if both predicted species occurrences and catch data were positive. Finally, the last criteria determined a spatial threshold to confirm that a species occurs in both neighboring EEZs. Once transboundary species were identified, we estimated their contribution to fishing landings and revenues and categorized them according to catch trends. We carry out the analysis at the species per EEZ level, rather than stocks (populations within a species), due to lack of stock-specific spatial and ecological information on all of the fished species [@Teh:2015gd]. Despite loosing detail-specific (e.g. a species could have multiple stocks within an EEZ), many fisheries within a country or EEZ are managed at the species level, and recent research has suggested strong connectivity between stocks thousand of kilometers away [@Ramesh:2019va].

@@

We identified **633** transboundary exploited species in the world, more than double the number previously estimated [@Teh:2015gd], accounting for **65.4%** of the species analyzed (Fig. 1A). Between 2005 and 2014, national fleets targeting shared species within EEZs landed an average of **69.2** million tonnes, representing **64%** of the EEZ-global catch, and generating USD **79,745** million in fishing revenue. This is equivalent to **49.8%** of the global revenue of the same period. These values are higher than previously estimated [@Teh:2015gd], reinforcing the importance of these species at the global level.

@@

![Fig. 1. Number of transboundary species and they contribution to fishing catch and revenue. A) The number of species and revenue are represented in the EEZ and land polygon, respectively. B) Contribution of tranbsoundary species to regional revenue (left) and catch (right). Points = mean +- sd.](./Figures/Fig1.png)

Transboundary species are particularly important for Northern America (mean = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (mean = 2,827 $\pm$ = 2,667 million USD), significantly differing from other regions in terms of revenue (one-way ANOVA; *F*(16,157) = 4.91, *p* < 0.001, $\alpha$ = 0.05) (Fig 1B). The large variation in revenue from these regions are driven by China (USD 6,279 million) in E. Asia and the US (USD 9,605 million) in N. America who contribute 55% and 83% to the region's revenue, respectively. Russia (USD 6,361 million), Peru (USD 5,211) and Japan (USD 3,369) complete the top five countries with more revenue from transboundary species. These five nations are responsible for 30% of the global revenue from transboundary species. Peru and Russia, responsible for the two largest fisheries in the world [@FAO:2018KK], lead the production of transboundary species harvesting over 5,000 thousand tones (**Supplemental figure Trans_Catch_MAP**). Peru's catches come mainly from Peruvian anchoveta (*Engraulis ringens*), shared with Chile, accounting for 79% of the transboundary species production. Similarly, Russia shares the Alaskan polllock (*Theragra chalcogramma*) fishery with both Japan and the US. Despite this, catches are only significantly different in E. Asia (one-way ANOVA; *F*(16,155) = 3.341, *p* < 0.001, $\alpha$ = 0.05), with China, the world's top fish producer with 15.3 million tons landed in 2016 [@FAO:2018KK] getting around 1/3 of its capture from transboundary species, followed by Japan (4.2 million tonnes) and South Korea (2.3 million tonnes).

Harvest rates and fisheries management are key to achieve the sustainable development goals [@SDGs]. We found significant differences in the exploitation rate of transboundary species (one-way ANOVA *F*(5,446), *f* = 324.5, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05 for all) with most species fished at levels below 50%, sometimes lower than 10%, of their historical maximum peak (Fig.2). This is the case of coastal nations in all continents, but Oceania, with no significant difference between continents (one-way ANOVA *F*(5,443), *f* = 1.24, *p* = 0.289, $\alpha$ = 0.05). While these catch trends do not directly reflect stock status - e.g. other factors like markets, fishing policies or extreme events also influence catch volume [@Pauly:2013gs; @Branch:2008kk] - they could resemble fisheries that are over exploited or collapsed [@Kleisner:2011wn]. Thus suggesting that the current global exploitation rate of transboundary species could be over the maximum sustainable yield [@Kleisner:2011wn; @Pauly:2013gs]. Corroborating the status of species harvested at this rate (Category C) and considering best practices in managing shared resources [@Munro:2004ug; @Lodge:2007bt] could provide crucial to the sustainability of transboundary species.

```{r Call_Circ_Country, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species by catch category, country, and continent. Category A, fishing between 10%-50% of max catch peak after a drop larger than 10% of max peak; Category B, fishing over 50% of max catch peak; Category C, fishing is between 10-50% of max catch peak or less than 10%"}

@@

![Fig. 2. Number of transboundary species by catch category. Category A, fishing between 10%-50% of max catch peak after a drop larger than 10% of max peak; Category B, fishing over 50% of max catch peak; Category C, fishing less than 50% of max peak](./Figures/Fig2.png)

Empirical analysis suggest that in most cases, management of transboundary species will have better outcomes in terms of fish yield when nations cooperate [@Bailey:2010ga]. Examples vary between regions and species including Mexico and the United States over Pacific anchovies [@CisnerosMontemayor:2020kv], Norway and Russia over Atlantic cod (*Gadus morhua*) [@Diekert:2010gp; @Eide:2013dha], and Namibia and South Africa over hake (*Merluccius spp*) [@Sumaila:2003vw]. Lack of collaboration in shared stocks may threaten stock sustainability [@Clark:1980js; @Nguyen:2018dk], reduce the profitability potential of the fishery [@Merino:2007jz], although cooperation and non-cooperation can be a more complex process [@Jensen:2015cf]. Collaboration can also reduce conflict between nations over shared resources. Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts like Canada, the US, China and Russia [@Spijkers:2019cz]. For example, the United States and Canada had a dispute over Pacific salmons (*Oncorhynchus sp*) that lasted over 10 years and was only resolved after these two nations set a series of regulations to compensate losses [@Munro:2002uf]. Since 2007, the European Union, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches (TACs) for Atlantic mackerel (*Scomber scombrus*) [@Spijkers:2017ij], a species that is transboundary and contributes **698.3** thousand tones to landings and USD **968** million in revenue to the countries involved in the dispute. In addition to past and present conflicts, climate change could exacerbate conflicts over transboundary species as climate change shifts the distribution of marine species around the globe requiring flexible agreements that accommodate future uncertainties [@Pinsky:2018cb].

@@

Over 130 marine fish species occur in tropical and subtropical waters around the world [@Nelson:2016kk], many of which are highly migratory pelagic-oceanic species like skipjack tuna (*Katsuwonus pelamis*), Common dolphinfish (*Coryphaena hippurus*), and tunas (*Thunnus sp.*). Pelagic-oceanic species are the only group whoes EEZ range (e.g. the number of EEZs where the species occur as transboundary) significantly differs from the others (one-way ANOVA *F*(5,109), *f* = 14.67, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 EEZs per species (Fig. 3).

@@

Due to their large migration patterns and presence in areas beyond national jurisdiction, highly migraroty pelagic fish species are cataloged as straddling and many of them managed by Regional Fisheries Management Organizations (RFMOs). Fifteen major RFMOs in the world oversee the management of 99 species [@CullisSuzuki:2010fi]. We estimate that 81.8% of these species are actually shared within neighboring RFMOs, with the International Commission for the Conservation of Atlantic Tunas sharing over 40 species with neighboring RFMOs (**Supplemental figure RFMO_MAP**). The effectiveness of RFMOs responsible for managing high migratory stocks has been previously questioned [@CullisSuzuki:2010fi; Lodge:2007bt]. However, despite a global expansion of the industrial fleet in the last decade [@Coulter:2020gr], a recent analysis of 13 species (40 stocks) suggests that most tuna and billfish stocks are sustainably harvested ($B/B_{MSY} > 1$ and $F/F_{MSY} < 1$) with only three stocks currently overfished [@Pons:2016ep]. Within national waters, we estimated that the catch of transboundary pelagic-oceanic species comes mainly from five Island Nations in the South Pacific from which three (Kiribalti, Solomon and Papua New Guinea) are members of the Pacific Nauru Agreement (Fig 3B). The parties to the Nauru Agreement sustainably manage 30% of global skipjack tuna [@Aqorau:2018bh], estimated to produce $2 billion worth of fish every year, representing a role model for transboundary fisheries management [@Hanich:2010ht].

@@

Is worth mention that for the current study we did not consider climate change as a driver of species' distribution. Dsitribution shifts are the most recorded response of marine species to the effects of climate change in the ocean [@Poloczanska:2016kk]. Moreover, catch composition might have already changed in the last five decades [@Cheung:2013gk], and will most likeley continue throuugh the 21^st century [@Cheung:2016jd]. While considering climate change in the distribution of shared stocks is beyond the scope of this study, it is likely that it will bring at least one new transboundary species per EEZ in the future [@Pinsky:2018cb], while also change the proportion of the distribution (e.g. Area Index) that each county's share to a degree that could influecne fisheries management [@Sumailaetal_in_press]. Exploring the impacts of climate change in share species is key to develop adaptation methods to minimize the effect of such distributional change on the conservation of species as well as the economy and food security of coastal nations.

Our study identifies the current species shared by neighbouring coastal nations and highlights the importance of these species in terms of global capture and revenue from wild fisheries. Our results show that captures and revenues from shared species are higher than previously estimated and particularly important to countries in Northern America and Eastern Asia. This result highlights the importance of these fisheries and their potential contribution to regional food security and wellbeing. Moreover, the vast majority of transboundary species are harvest at low rates with no geographic variation. Previous work has found that collaboration will result in better outcomes for shared marine resources. Thus, countries should pay specific attention in the management of the identified transboundary stocks potentially creating collaborative management plans. Identifying existing transboundary species is the first step towards joint management frameworks that are precautionary, strive for sustainability, and can be flexible to accommodate the uncertain future driven by climate change.

@@

All data is available by request of the authors.

@@

We determine whether or not a species can be considered transboundary. The method is based on a series of criteria to be met and the overlay of species distributions and the spatial boundaries of the world’s EEZs. For the current analysis we work at the species level and consider transboundary species as those that occur within the Exclusive Economic Zone (EEZ) of neighboring countries [@Munro:2002uf]. All data was scaled to $0.5^o$ x $0.5^o$ lat-long grid cells (*n* = ~180,000 grid cells).

## Databases on species geographic distributions

To determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs we used the list of commercial marine fish and invertebrate species from the *Sea Around Us* (SAU) database (http://www.seaaroundus.org) and determined their current distributions. For this, we used four data sources of species-distributions: (*i*) observational data, (*ii*) an Environmental Niche Model (ENMs), (*iii*) a life-history-based distribution model, and (*iv*) fisheries catch data (**Table x**). Each source represent a different method of estimating the distribution of a given species and thus, provides a more robust result. Only commercial fished species with data from all four sources were included in the analysis. The final dataset comprised 968 marine species (including andornomus species like salmons), **XX**% of the reported marine species in the SAU database and **XX**% of the reported taxa of the FAO (**See S1_Species_List**).

@@

The occurrence data was collected from five publicly available repositories; Fishbase (http://fishbase.org), the Global Biodiversity Information Facility (GBIF; https://www.gbif.org/), the Ocean Biogeographic Information System (OBIS; https://obis.org/),  the Intergovernmental Oceanographic Comission (IOC; http://ioc-unesco.org), and the International Union for Conservation of Nature (IUCN; https://www.iucn.org/technical-documents/spatial-data) [@Reygondeau:2019uh].

@@

In addition to the occurrence data we use two different methods to estimate species distributions, hereafter referred as ENM-Nereus and SDM-SAU. Although they use the same data, the models are structurally different complementing each other and providing robustness to the results.

The ENM-Nereus consists in a multimodel approach based on a Bioclim and a Bososted Regression Tree model [@Thuiller:2009gp], a Maxent model [@Phillips:2006ff], and a Non-Parametric Probabilistic Ecological Niche Model [@Beaugrand:2011fd]. Environmental variables utilized in the models included sea surface temperature, surface pH, surface oxygen concentration, and vertically integrated (0–100 m) net primary production (NPP) [@Asch:2018ca]. The ENM-Nereus employed all of the observational data-sources previously mentioned. Global environmental conditions were obtained from 1970 to 2000 and average from three Earth System Models developed by the Geophysical Fluid Dynamics Laboratory (GFDL- https://www.gfdl.noaa.gov/earth-system-model/), the Institute Pierre Simon Laplace (IPSL- www.icmc.ipsl.fr/), and the Max Planck Institute for Meteorology (MPI- www.mpimet.mpg.de/en/science/models/). See [@Asch:2018ca; @Reygondeau:2019uh] for model details.

The SDM-SAU model follows a five-steps process based on species-specific life history information, rather than environmental variables [@Close:2006ux; @Palomares:wg]. For each commercial fish species, the model first uses the FAO major fishing areas and countries EEZs to determine a broad distribution. It then uses life history information to delimit its range within the FAO fishing area (e.g. thermal preference, depth limits). The range is delimited even further by expert-review polygons and compared with that of AquaMaps [@Kaschner:2016tl], OBIS and GBIF occurrence data. The model then determines a species habitat preference based on the assumptions that the relative abundance of a species is determined by the number of habitats in a grid cell and the distance of the species to each habitat, as well as the importance of the habitat to the species size. Finally, the species equatorial submergence (e.g. the latitudinal region where a species is not seen in between poles) is estimated for each species. See [@Close:2006ux; @Palomares:wg] for model details.

@@

The previous models combine observational data with a series of biotic and abiotic information to determine the probability that a species will be found in a given space at a given time. However, this does not mean that the species in question will actually be there. While the models do use approaches to double-check species occurrences (e.g. ENM-Nereus uses four different species distribution algorithms and SPD-SAU undertakes validation by means of other models), we used a fourth data source to corroborate the models’ outputs. The *Sea Around Us* estimates total reconstructed catches - i.e., catches  based on all publicly available information sources and including discards, as well as unreported and illegal catches that are not included in the FAO data available for each country. Catches are also spatially allocated [@Zeller2016]. Thus, we used the Sea Around Us catch reconstruction database from 2005 to 2014 (last ten years of data) as the fourth dataset to estimate transboundary species and to estimate their catch contribution within EEZs.

@@

## Determine if a species is transboundary

We developed a four-criteria method for determining whether or not a species was transboundary. Only species that meet all criteria were considered as transboundary species and the analysis was done only within the EEZs of coastal states. In some cases, criteria also work as indexes to measure the uncertainty in the analysis as they provided continuous index (e.g. from 0 to 1) rather than a discrete, yes or no, index . All of the analysis was done in the statistical software *R version 3.5.2 (2018-12-20)* with the packages *data.table*[@Packagedatatable:2019uh], *janitor*[@PackagejanitorSim:2018ur], *wesanderson*[@Packagewesanderson:2018vn], *rfishbase*[@PackagerfishbaseR:2019th], *R.matlab*[@PackageRmatlabRe:2018un], *sf*[@PackagesfSimpleF:2018vp], *sp*[@PackagespClasses:2019tn], *tidiverse*[@PackagetidyverseE:2017vq], *tidytext*[@PackagetidytextTe:2019vk], and *zoo*[@PackagezooSInfr:2019tm]. All code is available at https://github.com/jepa/FishForVisa.

@@

We use the occurrence database, the ENM-Nereus model, and SDM-SAU model to determine the presence of each species within each of the 0.5 x 0.5 grid cells. We computed a *Species Index* by dividing the number of datasets that confirm presence of a species in a grid cell over the total amount of datasets (*n* = 3). The index is unit-less and represents a scale from 0.3 to 1 where 0.3 means only one dataset reports presence of the species in a grid cell, 0.6 represents two of three sources, and 1 means agreement between all datasets. For the analysis, we selected only those cases where *Species Index* = 1 to get a more conservative estimate of transboundary species.

@@

![**Number of shared species by RFMOs. Analysis included with 99 species managed by Fourteen RFMOs **](./Figures/S2_RFMO.png)

>>>>

Forty years after the formal adoption of UNCLOS, accurate estimates of the number of exploited marine species that are shared by neighboring nations (hereafter called transboundary species [@Munro:2002uf]) is still unclear. **Early data limitations allowed only for an informed guessed [@Caddy:1997ue], while recent estimations, followed a litterature review and catch data to address this issue [@Teh:2015gd]. However, up to date, no method has included the actual distribution of species. We hipothesize that the methodological constraints of preious work has underestimate the contribution of these species to global catch and revenue and that the proportoin of transboundary species is larger than previouslly estimated.** Thus, we aimed to estimate the number of exploited marine fish species shared by neighbouring countries and determine their contribution to global catch and fisheries revenue. **Moreover, we estimate their importance in terms of catch and revenue by country and also carry the analysis at the ecological level.**

We overlaid the known distribution of **968** commercially valuable marine species responsable for **%%** of global catch, and the EEZ of 153 coastal countries, and developed a four-criteria methodology to determine if a species was transboundary. First we only considered shared species between neighboring EEZs, rather than the species' extended distribution (e.g. we did not include the high seas). Secondly, we considered a species to be present in a grid cell if all data sources predicted positive occurrence (see Methods). The third criteria determined whether a species was exploited within a given grid cell using spatially explicit catch data. A species was considered as present in a grid cell if both predicted species occurrences and catch data were positive. Finally, we determined a spatial threshold based on the proportion of the species' distribution within each neighboring EEZs to confirm occurance in both EEZs.

Once transboundary species were identified, we estimated their contribution to fisheries catch and revenues and categorized them according to catch trends. We define a ‘stock’ unit as a species in an EEZ, instead of using a population ecology definition (genetically or morphological distinct units) due to lack of such biological information for almost all the species included in this study [@Teh:2015gd]. While we acknowledge that a species could have multiple stocks within an EEZ, many fisheries within a country or EEZ are managed at the species instead of stock level. Also, recent research has suggested strong connectivity between stocks, even when separated by thousand of kilometers, **and providing additional ecological ground for...** [@Ramesh:2019va].

@@

We identified **633** transboundary exploited species in the world, more than double of previous estimates [@Teh:2015gd], representing **65.4%** of the species analyzed (Fig. 1A). Between 2005 and 2014, national fleets targeting shared species within EEZs landed an average of **69.2** million tonnes, representing **64%** of the EEZ-derrived catches, and generating USD **79,745** million in fishing revenue. This is equivalent to **49.8%** of the global fishing revenue of the same period. These values are more than double the 34.2 million t and USD 30,700 million in fishing revenue previously estimated [@Teh:2015gd], underscoring the importance of transboundary species at the global level.

@@

![Fig. 1. Number of transboundary marine species and their contribution to global fisheries catches and revenue. A) The number of species and amount of revenue are represented by colour coding of EEZs and land polygons, respectively. B) Contribution of tranbsoundary species to regional revenue (left) and catch (right). Points = mean +- sd. Revenue in 2010 real USD](./Figures/Fig1.png)

**NOTE: EXPLORE THE REGIONAL SEGMENTATION YOU MADE**

Transboundary species are particularly important economically for Northern America (3,901 $\pm$ 4,497 million USD) and Eastern Asia (2,827 $\pm$ = 2,667 million USD), significantly differing from other regions in terms of revenue (one-way ANOVA; *F*(16,157) = 4.91, *p* < 0.001, $\alpha$ = 0.05) (Fig 1B). China (USD 6,279 million) in E. Asia and the US (USD 9,605 million) in N. America contribute 55% and 83% to their respective region's revenue, respectively. Russia (USD 6,361 million), Peru (USD 5,211) and Japan (USD 3,369) complete the top five countries with the most fishing revenue generated from transboundary species. These five nations are responsible for 30% of the fishing global revenue from transboundary species. Peru and Russia, responsible for the two largest fisheries in the world [@FAO:2018KK], lead transboundary catch statistics harvesting over 5,000 thousand tones each on average between 2005 and 2014 (**Supplemental figure Trans_Catch_MAP**). Peru's catches are mainly composed of Peruvian anchoveta (*Engraulis ringens*), shared with Chile, and accounting for 79% of the transboundary species production. Similarly, Russia shares the Alaskan polllock (*Theragra chalcogramma*) fishery with both Japan and the US. Despite the large catch volume of Peru and Russia, regional catches are only significantly different in E. Asia (one-way ANOVA; *F*(16,155) = 3.341, *p* < 0.001, $\alpha$ = 0.05), with China, the world's top fish producer with 15.3 million tons landed in 2016 [@FAO:2018KK] getting around 1/3 of its capture from transboundary species, followed by Japan (4.2 million tonnes) and South Korea (2.3 million tonnes). **NOTE: MAKE SURE RUSSIA IS NOT E.ASIA**

We determined the exploitation category of transboundary fisheries according to each country's catch [@Kleisner:2011wn] and also found significant differences among exploitation categories (one-way ANOVA *F*(5,446), *f* = 324.5, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05 for all) with most species fished to levels below 50%, sometimes lower than 10%, of their historical maximum catch (Fig.2). All regions conformed to this trend, except for Oceania, with no significant difference across continents (one-way ANOVA *F*(5,443), *f* = 1.24, *p* = 0.289, $\alpha$ = 0.05). While many environmental and social-economic factors such as temperature, markets, fishing policies affect catch volume [@Pauly:2013gs; @Branch:2008kk] previous studies demonstrated that the trends can be used to infer whether a stock is over-exploited or collapsed [@Kleisner:2011wn]. Thus suggesting that the current global exploitation rate of transboundary species could be over the maximum sustainable yield [@Kleisner:2011wn; @Pauly:2013gs].

**NOTE: THIS (L 118) PARAGRPAH NEEDS WORK**

```{r Call_Circ_Country, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species by catch category. Category A, fishing between 10%-50% of maximum catch level after a drop larger than 10% of the recorded maximum catch; Category B, fishing over 50% of maximum catch level; Category C, fishing less than 50% of maximum catch"}

@@

![Fig. 2. Number of transboundary species by catch category. Category A, fishing between 10%-50% of maximum catch level after a drop larger than 10% of the recorded maximum catch; Category B, fishing over 50% of maximum catch level; Category C, fishing less than 50% of maximum catch](./Figures/Fig2.png)

Empirical analysis suggests that in most cases, management of transboundary species will have better outcomes in terms of fish yields when nations cooperate [@Bailey:2010ga]. Examples of successful co-management include Canada and the United States over Pacific salmon [@REF] and Pacific halibut [@REF], Norway and Russia over Atlantic cod (*Gadus morhua*), haddock (*Melanogrammus aeglefinus*) and capelin (*Mallotus villosus*) in the Barents Sea [@Diekert:2010gp; @Eide:2013dha **NOTE: DOUBLE CHECK REFERENCES]**, and Namibia and South Africa over hake (*Merluccius spp*) [@Sumaila:2003vw]. Lack of collaboration over shared stocks may threaten stock sustainability [@Clark:1980js; @Nguyen:2018dk], and reduce the future profitability potential of the fishery [@Merino:2007jz]. Yet, we acknowledge that cooperation or lack thereof can be a complex process [@Jensen:2015cf] and that there are specific cases where cooperation might not be the best strategy [@REF]. Collaboration can also reduce conflict between nations over shared resources **as climate change shifts the distribution of species**. Transboundary fisheries are abundant and of economic importance for some of the top countries with past fisheries-related conflicts including Canada, the US, China and Russia [@Spijkers:2019cz]. For example, the United States and Canada had a dispute over Pacific salmons (*Oncorhynchus sp*) that lasted over 10 years and was only resolved after these two nations set a series of regulations were introduced to compensate for losses [@Munro:2002uf]. Since 2007, the European Union, Norway, Iceland, and the Faroe Islands have been at odds over the size and relative allocation of total allowable catches (TACs) for Atlantic mackerel (*Scomber scombrus*) [@Spijkers:2017ij], a species that is transboundary and contributes **698.3** (**NOTE: %%**) thousand tonnes to catches and USD **968** (**NOTE: %%**) million in revenue to the countries involved in the dispute. In addition to past and present conflicts, climate change is likely to exacerbate conflicts over transboundary species due to shifts in their distribution, requiring flexible agreements that accommodate future uncertainties [@Pinsky:2018cb].

@@

Over 130 marine fish species occur in tropical and subtropical waters around the world [@Nelson:2016kk], many of which are highly migratory pelagic-oceanic species like skipjack tuna (*Katsuwonus pelamis*), other tunas (*Thunnus sp.*), sword fish (*Xiphias gladius*), and common dolphinfish (*Coryphaena hippurus*). Pelagic-oceanic species are the only group whose EEZ range (i.e., the number of EEZs where the species occur as 'transboundary') was significantly different from other species (one-way ANOVA *F*(5,109), *f* = 14.67, *p* < 0.001, $\alpha$ = 0.05; Tukey's post hoc test *p* < 0.05) with a median of 40 EEZs per species (Fig. 3).

@@

Due to their large migration patterns and presence in areas beyond national jurisdiction, many highly migraroty pelagic fish species are managed by Regional Fisheries Management Organizations (RFMOs). Fifteen major RFMOs in the world oversee the management of 99 species [@CullisSuzuki:2010fi]. We estimated that 81.8% of these species are shared across neighboring RFMOs, with the International Commission for the Conservation of Atlantic Tunas (ICCAT) sharing over 40 species with neighboring RFMOs (**Supplemental figure RFMO_MAP**). The effectiveness of RFMOs responsible for managing high migratory stocks has been previously questioned [@CullisSuzuki:2010fi; @Lodge:2007bt]. However, despite a global expansion of the industrial fleet in the last decade [@Coulter:2020gr], a recent analysis of 13 species (40 stocks) suggests that most tuna and billfish stocks are sustainably harvested ($B/B_{MSY} > 1$ and $F/F_{MSY} < 1$) with only three stocks currently considered as overfished ($B/B_{MSY} < 1$ and $F/F_{MSY} > 1$) [@Pons:2016ep]. Within national waters, we estimated that the catch of transboundary pelagic-oceanic species mainly originates from five large ocean states in the South Pacific, three of which (Kiribati, Solomon Islands and Papua New Guinea) are parties to the Pacific Nauru Agreement (Fig 3B). Parties to the Nauru Agreement sustainably manage 30% of global skipjack tuna [@Aqorau:2018bh], estimated to produce $2 billion worth of fish every year, representing a role model for transboundary fisheries management [@Hanich:2010ht].

**NOTE: Colette will search for a reference related a weird comment "arecenWCPF assessment highlight tt tuna in the Pacific e sustsinabl fisd, but oher species  socks in ot regionsre . I will dig up e reference and send it to you - just to make sure  the irmati overlaps"**

@@

In our analyses we did not seek to include climate change as a driver of species’ distribution. Distribution shifts are the most commonly recorded response of marine species to the effects of climate change - chiefly warming - in the ocean [@Poloczanska:2016kk]. Catch composition might have already changed in the last five decades [@Cheung:2013gk], and will continue throuugh the $21^{st}$ century [@Cheung:2016jd]. Previous study projected that climate change will bring at least one new transboundary species per EEZ in the future [@Pinsky:2018cb], while also change the proportion of the distribution (e.g. Area Index) that each county’s share to a degree that could influence fisheries management (*Palacios-Abrantes et.al. under review*). Exploring the detailed impacts of climate change in share species is key to develop local adaptation methods to minimize the effect of such distributional change on the conservation of species as well as the economy and food security of coastal nations.

Our study identified species currently shared by neighbouring coastal nations and highlights the importance of these species' contribution to global capture and revenue from wild fisheries. Our results show that captures and revenues from shared species are higher than previously estimated and particularly important to countries in Northern America and Eastern Asia. This result highlights the importance of these fisheries and their potential contribution to regional food security and wellbeing. Moreover, the vast majority of transboundary species are harvest at low rates with no geographic variation. Previous work has found that collaboration will result in better outcomes for shared marine resources. Thus, countries should pay specific attention in the management of the identified transboundary stocks potentially creating collaborative management plans. Identifying existing transboundary species is the first step towards joint management frameworks that are precautionary, strive for sustainability, and can be flexible to accommodate the uncertain future driven by climate change.

@@

All data is available from the authors upon request.

@@

To ascertain whether or not a species can be considered transboundary, we developed a method based on a series of criteria and the overlay of species' distributions and the spatial boundaries of the world’s EEZs. All analyses were conducted at the species level and we consider transboundary species as those that occur within the Exclusive Economic Zone (EEZ) of neighboring countries [@Munro:2002uf]. All data was scaled to $0.5^o$ x $0.5^o$ lat-long grid cells.

## Databases of species' geographic distribution

To determine the number of transboundary marine species exploited by fisheries within each of the world's EEZs we used the list of commercial marine fish and invertebrate species from the *Sea Around Us* (SAU) database (http://www.seaaroundus.org) and determined their distributions. To this end, we used four data sources of species-distributions: (*i*) occurence data, (*ii*) an Ensemble Environmental Niche Model (ENMs), (*iii*) a life-history-based distribution model, and (*iv*) fisheries catch data (**Table x**). Each source represents a different method of estimating the distribution of a given species and thus, provides a more robust result. Only commercial fished species with data from all four sources were included in the analysis **NOTE: WC "Need to discuss how this affect the samples of transboundary species."**. The final dataset comprised 968 marine species (including anadromous species like salmononids), **XX%** of the total number of reported marine species in the SAU database and **XX%** of the reported marine taxa of included by the Food and Agricultural Organization of the United Nations (FAO).

<!-- **NOTE: for transboundary lobsters https://books.google.com.ar/books?id=XEElAwNrKZsC&pg=PA224&lpg=PA224&dq=belize+mexico+lobster+shared+stock&source=bl&ots=lsmUH743d_&sig=ACfU3U1P1t7_L8q_YRh3oQGszUO8Lg1jFw&hl=en&sa=X&ved=2ahUKEwiv36qbwbjnAhX3HrkGHYIbCTgQ6AEwFHoECA0QAQ#v=onepage&q=belize%20mexico%20lobster%20shared%20stock&f=false** -->

@@

Occurrence data was collected from five publicly available repositories: Fishbase (http://fishbase.org), the Global Biodiversity Information Facility (GBIF; https://www.gbif.org/), the Ocean Biogeographic Information System (OBIS; https://obis.org/),  the Intergovernmental Oceanographic Comission (IOC; http://ioc-unesco.org), and the International Union for Conservation of Nature (IUCN; https://www.iucn.org/technical-documents/spatial-data) [@Reygondeau:2019uh].

@@

In addition to occurrence data, we used two different methods to estimate species distributions, hereafter referred to as Ecological Niche Model-Nereus (ENM-Nereus) and Species Distribution Model-SAU (SDM-SAU). Although they use the same occurrence and environmental data, the models are structurally different complementing each other and providing robustness to the results.

The ENM-Nereus consists in a multimodel approach based on a Bioclim and a Boosted Regression Tree model [@Thuiller:2009gp], a Maxent model [@Phillips:2006ff], and a Non-Parametric Probabilistic Ecological Niche Model [@Beaugrand:2011fd]. Environmental variables utilized in the models include sea surface temperature, surface pH, surface oxygen concentration, and vertically integrated (0–100 m) net primary production (NPP) [@Asch:2018ca]. The ENM-Nereus employed all of the observational data-sources previously mentioned. Global environmental conditions were average for the period 1970-2000 and averaged for three Earth System Models developed by the Geophysical Fluid Dynamics Laboratory (GFDL- https://www.gfdl.noaa.gov/earth-system-model/), the Institute Pierre Simon Laplace (IPSL- www.icmc.ipsl.fr/), and the Max Planck Institute for Meteorology (MPI- www.mpimet.mpg.de/en/science/models/). See [@Asch:2018ca; @Reygondeau:2019uh] for model details.

The SDM-SAU model follows a five-steps process based on species-specific life history data, rather than environmental variables [@Close:2006ux; @Palomares:wg]. For each commercial species, the model first uses the FAO major fishing areas and countries' EEZs to determine a broad distribution. It then uses life history information to delimit its range within the FAO fishing area (e.g., thermal preference, depth limit). The range is delimited even further by expert-review polygons and compared with that of AquaMaps [@Kaschner:2016tl], OBIS and GBIF occurrence data. The model then determines a species habitat preference based on the assumption that the relative abundance of a species is determined by the number of habitats in a grid cell and the distance of the species to each habitat, as well as the importance of the habitat to the species size. Finally, the species equatorial submergence (e.g. the latitudinal region where a species is not seen in between poles) is estimated for each species. See [@Close:2006ux; @Palomares:wg] for model details.

@@

The previous models combine observational data with a series of biotic and abiotic information to determine the probability that a species will be found in a given space at a given time. However, this does not mean that the species in question will actually be there. While the models do use approaches to double-check species occurrences (e.g., ENM-Nereus uses four different species distribution algorithms and SPD-SAU undertakes validation by means of other models), we used a fourth data source to corroborate the models’ outputs. The Sea Around Us estimates total reconstructed catches - catches based on all publicly available information sources and including discards, as well as unreported and illegal catches that are not included in available FAO data - for each country. Catches are also spatially allocated [@Zeller2016]. Thus, we used the Sea Around Us catch reconstruction database from 2005 to 2014 as the fourth dataset to estimate transboundary species and to estimate their catch contribution within EEZs. **We selected this time frame to investigate the recent (our time frame includes the last ten years of data) contribution of transboundary fisheries and to reduce the uncertanty embeded in the recosntruction process (See Key uncertanties).**

@@

## Determining transboundary species trait

We developed a four-criteria methodology to determine whether or not a species can be considered transboundary. Only species that meet all criteria were considered as transboundary and the analysis was undertaken only within the EEZs of coastal states. In some cases, criteria can also be interpreted as indices to measure the uncertainty in the analysis as they provide a continuous index (e.g. from 0 to 1) rather than a discrete, yes or no, index. All analyses were run using the statistical software *R version 3.5.2 (2018-12-20)* with the packages *data.table*[@Packagedatatable:2019uh], *janitor*[@PackagejanitorSim:2018ur], *wesanderson*[@Packagewesanderson:2018vn], *rfishbase*[@PackagerfishbaseR:2019th], *R.matlab*[@PackageRmatlabRe:2018un], *sf*[@PackagesfSimpleF:2018vp], *sp*[@PackagespClasses:2019tn], *tidiverse*[@PackagetidyverseE:2017vq], *tidytext*[@PackagetidytextTe:2019vk], and *zoo*[@PackagezooSInfr:2019tm]. All code is available at https://github.com/jepa/FishForVisa.

@@

We used the occurrence database, the ENM-Nereus model, and SDM-SAU model to determine the presence of each species within each of the $0.5^{\circ}$ x $0.5^{\circ}$ grid cells. We computed a *Species Index* by dividing the number of datasets that confirmed the presence of a species in a grid cell over the total number of datasets (n = 3). The index is unit-less and represents a scale from 0.3 to 1, with 0.3 indicating that only one dataset reported the presence of the species in a grid cell, 0.6 two out of three sources, and 1 pointing to agreement across all 3 datasets. All analyses only considered cases where *Species Index* = 1 to obtain a more conservative estimate of transboundary species.

@@

![**Number of shared species across RFMOs. Analysis included with 99 species managed by Fourteen RFMOs **](./Figures/S2_RFMO.png)
